### 2021-07-21 chapters/3_background.tex
move model description to background

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
Relevant models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain and subtractive clustering.
    \item Neural gas.
    \item ART networks.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

>>>>

\section{Machine learning models for clustering}
\label{ch:neuralNetworkSurvey}
There are many different types of machine learning models with their respective applications.
For this thesis we will focus on machine learning models designed for clustering.
\newline
Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
\newline
K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
Relevant models he discusses are:
\begin{itemize}
    \item Self organizing map.
    \item Learning vector quantization.
    \item C-means(K-means) clustering.
    \item Mountain and subtractive clustering.
    The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
\newline
The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points. Each grid point is a potential cluster center. 
The potential for each grid is calculated based on the density of the surrounding data points. 
The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
    \item Neural gas.
    The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
Unlike the \acrlong{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
A benefit of the neural gas network is that it is able to spawn neurons.
This means that it is not necessary to know the number of clusters in advance.
    \item ART networks.
    ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
    \item Fuzzy c-means.
\end{itemize}
K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
\newline
He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
Machine learning model variants and their references are also discussed.
\newline
He closes his paper with computer simulations of some of the machine learning models he surveyed.

### 2021-07-21 chapters/rq2/1_clustering_mutants.tex
move model description to background

\subsection{Mountain and subtractive clustering}
The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
\newline
The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points. Each grid point is a potential cluster center. The potential for each grid is calculated based on the density of the surrounding data points. The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
\newline
Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
If it is required to find out the number of clusters we need for our model we can make use of these models.

\subsection{ART models}
ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
While the ART models are good at what they do, they do not fit our purpose of clustering mutants.
As described above the ART models expect rapidly changing sequences of input patterns.
Our data does not contain binary input patterns and it is also not rapidly changing.
Therefore we do not select an ART model.

\subsection{Neural gas}
The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
Unlike the \acrlong{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
A benefit of the neural gas network is that it is able to spawn neurons.
This means that it is not necessary to know the number of clusters in advance.
Unfortunately the neural gas network is a learning network.
One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.
For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
Therefore we do not select the neural gas network.

>>>>

\subsection{Mountain and subtractive clustering}
As described in Chapter \ref{ch:neuralNetworkSurvey} the mountain and subtractive clustering models calculate the amount of clusters needed for a specific data set.
Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
If it is required to find out the number of clusters we need for our model we can make use of these models.

\subsection{ART models}
While the ART models are good at what they do, they do not fit our purpose of clustering mutants.
As described in Chapter \ref{ch:neuralNetworkSurvey} the ART models expect rapidly changing sequences of input patterns.
Our data does not contain binary input patterns and it is also not rapidly changing.
Therefore we do not select an ART model.

\subsection{Neural gas}
The neural gas network is a learning network.
One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.
For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
Therefore we do not select the neural gas network.
